---
title: "Yelp data Exploratory analysis"
author: "Sean Clarke"
date: "7 November 2015"
output: html_document
---


```{R cache=TRUE}
load("/Users/Sean/Coursera_DataScience/JHU-DataScience-Capstone/yelp_dataset_challenge_academic_dataset/yelpdata.rda")
```

```{r}
library(ggplot2)
library(mosaic)
```

Look for NAs

```{r}
colSums(is.na(yelpdata))
```

Is the stars coulmn clean

```{r}
unique(yelpdata$stars)
```

[THis is useful Useful!!!](https://www.yelp.com/developers/documentation/v2/all_category_list)

Find the top categories

```{r}
z <- data.frame(table(unlist(head(yelpdata$categories,n=1000000))))    # Count categories
topCats <- head(z[order(z$Freq, decreasing=T), ], n = 10)               # List cats in order

ggplot(data=topCats, aes(x=Var1, y=Freq)) +
  geom_bar(stat="identity")  + xlab("Category") + ylab("Number of Occurences") + theme(axis.text.x=element_text(angle=90, size=10, vjust=0.5))
```

Lets concentrate on restaurants

```{r}
yelpdata <- yelpdata[grepl( "^.*[Rr]estaurant.*",  yelpdata$categories),]
```

How does the geographical distribution look?


```{r}
unique(yelpdata$state)
#[1] "PA"  "NC"  "SC"  "WI"  "IL"  "AZ"  "NV"  "QC"  "ON"  "EDH" "MLN" "FIF" "ELN" "XGL" "BW"  "RP"  "KHL" "NW" 
```


WHere do those weird states map to?

PA      US - Philadelphia
NC      US - North Carolina
SC      US - South Carolina
WI      US - Wisconsin
IL      US - Illinois
AZ      US - Arizona
NV      US - Nevada
QC      Canada - Quebec
ON      Canada - Ontario
EDH     UK - Edinburgh
MLN     UK - Midlothian
FIF     UK - Fife
ELN     UK - Elgin? (Musselburgh)
XGL     UK - Edinburgh
BW      Germany - Ettlingen
RP      Germany
KHL     UK - Edinburgh
NW      Germany

```{r}
yelpdata <- mutate(yelpdata, country= derivedFactor(
               "DE" = state %in% c("NW", "RP", "BW"),
               "UK" = state %in% c("EDH", "MLN", "FIF", "ELN", "XGL"),
               "CA" = state %in% c("QC", "ON"),
                 .method = "first",
                 .default = "US"
             )
)

yelpdata %>%
  group_by(country) %>%
  summarise(length(country))
```
